---
title: "Exploring nutrition status- data wrangling"
author: "Eric Anku"
format: html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r import-libraries}
# Import libraries
pacman::p_load(
        tidyverse, # Data wrangling and plotting
        googlesheets4, # Import dataset
        anthro, # Prevalence and z-scores
        gtsummary, # Table summaries
        ggstatsplot, # Exploratory plots
        GGally, # Scatterplot matrix
        statsr, # Inferential statistics
        skimr, # Statistical summary
        finalfit, # Labelling, glimpse, and modelling
        janitor, # Data cleaning
        broom, # Tidy data output
        rstatix, # Normality test
        here, # Directory management
        rio # Import and export

)
```


## Import dataset

```{r import-dataset}
child <- read_sheet("https://docs.google.com/spreadsheets/d/1yL53EdTsd5fbDmAHltv79TFbVULi4KfxYXmRXA2njYA/edit?usp=sharing",
           col_names = c(
                   "meta_data",
                   "dob",
                   "dov",
                   "age",
                   "name",
                   "sex",
                   "bwt",
                   "weight",
                   "lenhei"
           ),
           col_types = "?DDdccddd",
           skip = 1)
```



## Assess dataset

- Assess for duplicates
- Assess for missingness
- Data types
- Unusual birth weight value {Check bwt: 304.0 kg}
- Unusual weight value {Check weight: 95 kg}
- Log transform numeric variables and assess normality


```{r head-dataset}
head(child)
```

```{r get-duplicates}
child %>% 
        get_dupes(name) # Assess duplicate names
```


```{r assess-weight-values}
# Assess weight values for unusual values
child%>%select(c(name, weight, dov))%>%filter(weight > 10)
```

```{r assess-birthweight-values}
# Assess birthweight values for unusual values
child%>%select(c(name, bwt, dov))%>%filter(bwt > 5)
```


```{r assess-lenhei-values}
# Assess length/height values for unusual values
child%>%select(c(name, lenhei, dov))%>%filter(lenhei > 95)

## There are no issues with the length/height data
```


```{r assess-age-values}
child%>%select(name, dov, age)%>%filter(age < 42)
```


```{r replace-wrong-birthweight-values}
# Child 1 
child$bwt[child$bwt == 42] <- 3.2

# Child 2
child$bwt[child$bwt == 182.6253] <- 4.0
 
# Child 3
child$bwt[child$bwt == 12.9] <- 2.9
 
# Child 4
child$bwt[child$bwt == 304] <- 3.2

# Child 5
child$bwt[child$bwt == 6.6] <- 3.6

# Child 6
child$bwt[child$bwt == 6] <- 3.0
```

```{r replace-wrong-weight-values}
# Child 1 
child$weight[child$weight == 86] <- 10.9

# Child 2
child$weight[child$weight == 70] <- 7.8
 
# Child 3
child$weight[child$weight == 95] <- 9.5
 
# Child 4
child$weight[child$weight == 64] <- 6.4

# Child 5
child$weight[child$weight == 42] <- 4.2

# Child 6
child$weight[child$weight == 55] <- 5.5

# Child 7
child$weight[child$weight == 50] <- 5.0

# Child 8
child$weight[child$weight == 59] <- 5.9

# Child 9
child$weight[child$weight == 2.3] <- 4.8

# Child 10
child$weight[child$weight == 1.60] <- 2.7
```


```{r replace-wrong-age-values}
# Child 1 
child$age[child$age == 14] <- 60.85

# Child 2
child$age[child$age == 35] <- 152.125
 
# Child 3
child$age[child$age == 14] <- 60.85
 
# Child 4
child$age[child$age == 35] <- 152.125

# Child 5
child$age[child$age == 28] <- 91.275

# Child 6
child$age[child$age == 30.4375] <- 42

# Child 7
child$age[child$age == 31.3125] <- 91.275

# Child 8
child$age[child$age == 30.4375] <- 42

# Child 8
child$age[child$age == 30.4375] <- 42

# Child 9
child$age[child$age == 38.4375] <- 42

# Child 10
child$age[child$age == 30.4375] <- 42

# Child 11
child$age[child$age == 30.4375] <- 42

# Child 12
child$age[child$age == 3.2] <- 42

# Child 13
child$age[child$age == 30.4375] <- 42
```

```{r replace-wrong-lenhei-values}
# Child 1 
child$lenhei[child$lenhei == 10.9] <- 86

# Child 2
child$lenhei[child$lenhei == 7.8] <- 78
 
# Child 3
child$lenhei[child$lenhei == 5.1] <- 51
 
# Child 4
child$lenhei[child$lenhei == 5.2] <- 52

# Child 5
child$lenhei[child$lenhei == 8.3] <- 83

# Child 6
child$lenhei[child$lenhei == 7.2] <- 72

# Child 7
child$lenhei[child$lenhei == 4.1] <- 41
```


```{r remove-duplicates}
# Remove duplicate entries
child <- child%>%
        distinct(name, .keep_all = TRUE)
```


```{r missing-data}
# Assess missing data
ff_glimpse(child)
```


```{r final-dataset}
child_final <- child%>%
        select(-c(
                meta_data,
                dob,
                dov
        )) # De-identify dataset and remove variables not needed
```


## Clean dataset

- Check for flagged children and verify their entries
- Correct mistake entries 
- Remove unusual weight, birthweight, length or height values
- Log transform numeric variables to minimise skewness and verify normality

```{r wrangle}
child_final <- child_final%>%
        mutate(
                age = age%>%
                        ff_label ("Age in days"),
                name = ff_label(name, "Name"),
                sex = sex%>%
                        ff_label("Sex"),
                bwt = bwt%>%
                        ff_label("Birth weight (kg)"),
                weight = weight%>%
                        ff_label("Weight (kg)"),
                lenhei = lenhei%>%
                        ff_label("Length/Height (cm)"),
                age_months = age/30.4375%>%
                  ff_label("Age in months")
        )
```

```{r export-data-csv}
export(child_final, here("data","cwc.csv"))
```

```{r export-data-rdata}
export(child_final, here("data","cwc.RData"))
```

